---
title: "Probability and Statistics"
subtitle: "The Evolution of Computer Processors: A statistic on Common Properties"
author: "Chau Dang Minh - 2013748 <br> Provide your name guys"
output: 
  # pdf_document: default
  html_document: default
---

```{r message=FALSE}
# Libraries and options
library(dplyr)
library(here)
library(knitr)
library(kableExtra)

# Self-defined functions
source("utils.R")

# Working directory
setwd(here())
```

# Dataset inspection

```{r}
# Read the CSV file into a data frame
cpu_data <- read.csv("dataset/Intel_CPUs.csv")
gpu_data <- read.csv("dataset/All_GPUs.csv")
```

Get the formatted first rows of the data

```{r}
# Inspect the CPU data
kable(head(cpu_data), format = "html") %>%
  kable_styling()
```

Take a look at the columns.

```{r}
cpu_columns <- colnames(cpu_data)
cpu_columns
```

```{r}
cpu_collections <- unique(cpu_data$Product_Collection)
length(cpu_collections)
```

There are 75 distinct values of <code>Product_Collection</code>. We will find out the key of this data.

```{r}
difference <- nrow(cpu_data) - nrow(unique(cpu_data[, c("Product_Collection", "Vertical_Segment", "Processor_Number", "Launch_Date")]))
difference
```

Let us check that the key of this data is the tuple <code>(Product_Collection, Processor_Number)</code>.

# Statistics

## How $\texttt{Recommended_Customer_Price}$ depends on other manufacturing properties

-   Filter out instances with no $\texttt{Recommended_Customer_Price}$
-   Choose concerned properties

```{r}
selected_cpu_data <- cpu_data[, c("Recommended_Customer_Price", 
  "nb_of_Cores", 
  "nb_of_Threads", 
  "Processor_Base_Frequency",
  "Bus_Speed")]
processed_cpu_data <- 
  selected_cpu_data[
    apply(selected_cpu_data, 1, function(row) all(sapply(row, is_valid))), ]
```